---
title: "GMSE European hamster model"
author: "Imke Tomsin"
date: "2023-2024"
output:
  word_document: default
  html_document: default
---

# Defining the different months that will be simulated

```{r, , echo=TRUE, eval=FALSE}
TMAX <- 60; # Maximum number of months in the simulation
JAN  <- seq(from = 1,  to = TMAX, by = 12); # Time steps in JAN, FEB, MAR, etc.
FEB  <- seq(from = 2,  to = TMAX, by = 12);
MAR  <- seq(from = 3,  to = TMAX, by = 12);
APR  <- seq(from = 4,  to = TMAX, by = 12);
MAY  <- seq(from = 5,  to = TMAX, by = 12);
JUN  <- seq(from = 6,  to = TMAX, by = 12);
JUL  <- seq(from = 7,  to = TMAX, by = 12);
AUG  <- seq(from = 8,  to = TMAX, by = 12);
SEP  <- seq(from = 9,  to = TMAX, by = 12);
OCT  <- seq(from = 10, to = TMAX, by = 12);
NOV  <- seq(from = 11, to = TMAX, by = 12);
DEC  <- seq(from = 12, to = TMAX, by = 12)
# This way, we are able to use each time step in the GMSE loop to identify the month
```

# Function to simulate carrying capacity by cell 

```{r, echo=TRUE, eval=FALSE}
cell_K <- function(res, DIM_1 = 447, DIM_2 = 447){
  for(i in 1:DIM_1){
    for(j in 1:DIM_2){
      sum_ij <- sum(res[, 5] == i & res[, 6] == j)
      if(sum_ij > 1){
        on_ij             <- which(res[, 5] == i & res[, 6] == j);
        loser_ij          <- sample(x = on_ij, size = sum_ij - 1);
        res[loser_ij, 7]  <- 0; # Stop in tracks
        res[loser_ij, 9]  <- 1; # Definite death
        res[loser_ij, 10] <- 0; # No birth
        res[loser_ij, 11] <- 0; # No birth
      }
    }
  }
  return(res);
}

#Now, if two hamsters occupy the same cell, one of them will die. 

#This should be run after movement has happened, so that offspring are not removed before
#reproductively active/old enough to leave the mother
#This function is quite slow -- probably want to speed it up somehow
```

# Function to add a crop type to each cell of the landscape 

```{r, , echo=TRUE, eval=FALSE}
# The 'land' in GMSE is a 3D array with the farm number in layer 3, the crop amount in layer 2, and nothing in layer 1. The code below redefines layer 1 by putting a crop type (1, 2 or 3) down for each farm number (you can change these as desired).The code assigns a random crop type to the land owned by each stakeholder. The function has to be run before looping over time. 

crop_type <- function(land, agents = 11, types = 3){ # agents = manager + number of stakeholders, types = number of possible crop types
  xdim <- dim(land)[1];
  ydim <- dim(land)[2];
  for(i in 1:xdim){   
  for(j in 1:ydim){
        if (land[i, j, 3] %in% 1:agents) {  #1:number of farms/stakeholders 
          land[i, j, 1] <- sample(1:types, 1) #1:number of possible crop types, each cell gets a different one with this function
      }
    }
  } 
    return(land);
} 

```

# Alternative functions to assign a fixed number of (random) stakeholders/farms a certain crop type

To assign a certain crop type to a certain amount of fields, and assign the other fields a random type:

```{r, , echo=TRUE, eval=FALSE}
crop_type <- function(land, agents = 11, types = 2) { #types = amount of crop types of which no fixed number is assigned to fields
  xdim <- dim(land)[1]
  ydim <- dim(land)[2]
  
  type3_stakeholders <- sample(2:agents, 5) # Select 5 random stakeholders to get assigned crop type 3
  
  # Create a vector to store the crop type for each stakeholder
  stakeholder_types <- rep(NA, agents)
  
  for (stakeholder in 2:agents) {
    if (stakeholder %in% type3_stakeholders) {
      stakeholder_types[stakeholder] <- 3
    } else {
      stakeholder_types[stakeholder] <- sample(1:types, 1)
    }
  }
  
  for (i in 1:xdim) {   
    for (j in 1:ydim) {
      if (land[i, j, 3] %in% 2:agents) {
        land[i, j, 1] <- stakeholder_types[land[i, j, 3]]
      }
    } 
  }
  return(land)
}
```

To assign a certain number of fields a certain crop type
```{r, , echo=TRUE, eval=FALSE}
crop_type2 <- function(land, agents = 10) { # agents = 1 (manager) + number of stakeholders
  xdim <- dim(land)[1]
  ydim <- dim(land)[2]
  
  # Randomly assign stakeholders to crop types
  stakeholders_vector <- sample(2:agents) #randomly shuffle the stakeholders (agents), but leave out the first (= always the manager)
  type3_stakeholders <- stakeholders_vector[1:2] #2 get type 3
  type2_stakeholders <- stakeholders_vector[3:6] #4 get type 2
  type1_stakeholders <- stakeholders_vector[7:10] #4 get type 1
  
  # Create a vector to store the crop type for each stakeholder
  stakeholder_types <- rep(NA, agents)
  
  for (stakeholder in 2:agents) { 
    if (stakeholder %in% type3_stakeholders) {
      stakeholder_types[stakeholder] <- 3
    } else if (stakeholder %in% type2_stakeholders) {
      stakeholder_types[stakeholder] <- 2
    } else if (stakeholder %in% type1_stakeholders) {
      stakeholder_types[stakeholder] <- 1
    }
  }

  for (i in 1:xdim) {   
    for (j in 1:ydim) {
      if (land[i, j, 3] %in% 2:agents) {
        land[i, j, 1] <- stakeholder_types[land[i, j, 3]]
      }
    } 
  }
  return(land)
}

```

# Functions to change population parameters based on crop type

**Crop type 1**: crop unsuitable for hamsters.
These are crops in which hamsters do not tend to occur at all, e.g. because of lack of food or coverage. In the model, hamsters have a survival probability of 0 when they establish in these fields. 

**Crop type 2**: intensive agriculture of a hamster-friendly crop.
These are crops that are consumed by hamsters, and in which they can survive. But, they are monocultures, and no hamster-friendly adjustments have been made to the way they are cultivated. Example: cereals that are harvested too early, leading to 0% coverage while the hamsters are still reproducing and not yet hibernating.

**Crop type 3**: Hamster-friendly crops, with adjustments for hamster-friendly agricultural practices.
Later harvests allow for an additional litter

Not in here yet but I might want to add it, crop type 4: Hamster heaven. 
No or very late harvest, with vegetation being a mix of hamster-friendly crops and herbs etc. leading to a higher lambda. 

```{r, , echo=TRUE, eval=FALSE}
# The function below checks each hamster's cell location, then checks the crop type on that location, and adjusts population parameters accordingly

crop_hamster <- function(land, res){     
    N <- dim(res)[1]; # Number of hamsters
    for(i in 1:N) { 
        xloc <- res[i, 5] +1; #+1 to fix indexing difference problem between R and C (otherwise error occurs about the landscape starting in (0,0))
        yloc <- res[i, 6] +1;
        crop <- land[xloc, yloc, 1];
        
        
         #cat("Hamster ", i, " at (", xloc, ", ", yloc, ") has crop type: ", crop, "\n") #added this bit to check if the function is working properly, this shows that some hamsters' xloc or yloc randomly seems to change by 1 cell between simulations, even when movement is not simulated, sometimes both change, sometimes just one changes, and sometimes the crop type value (1, 2 or 3) also changes
      
          if (crop == 1) {
          res[i, 9] <- 1  # column 9 = death probability -> 100% mortality
        #} else if (crop == 2 && next_time %in% c(AUG,SEP)) {
          #res[i, 9] <- res[i, 9] + 0.05  # increase mortality by 5% in months after harvest 
        } else if (crop == 3 && next_time %in% c(JUN, JUL, AUG)) { 
          res[i, 10] <- 3
        } else {
          # Default case, nothing changes
      }
    }
    return(res);
}

```

# GMSE simulations 

## 1: Simulation for a certain set of parameters, no changes are made to the landscape. 

2 litters, in June and July, movement in June and July, starting X of 15, 30, and 250 females (last one based on 50/500 rule)
(no changes to landscape/only one crop type at a time)

```{r, , echo=TRUE, eval=FALSE}

library(GMSE)
DIM_1 <- 447; # Land dimension 1
DIM_2 <- 447; # Land dimension 2

iterations <- 100 #number of iterations
results <- list() #list to store results from each repetition

for (rep in 1:iterations) {
  tryCatch({

# Initialise the first output
sim_old   <- gmse_apply(stakeholders    = 1, # not important here as landscape is all the same and no actions are being taken
                        res_movement    = 100, 
                        remove_pr       = 1 - 0.976083968, 
                        lambda          = 0, 
                        res_death_type  = 1,
                        observe_type    = 2, # but only once a year
                        res_move_obs    = FALSE, 
                        max_ages        = 24, 
                        RESOURCE_ini    = 250, # N 
                        culling         = FALSE,
                        land_ownership  = TRUE,
                        age_repr        = 4,
                        land_dim_1      = DIM_1,
                        land_dim_2      = DIM_2,
                        manage_target   = 2500,
                        res_move_type   = 0,
                        get_res         = 'Full');
old_obs  <- sim_old[["observation_array"]];

# Initialise the results matrix for the current iteration; Currently, 2 variables are recorded as output
# Double-brackets: technically a bit more secure. It works if you edit columns of the resource_array directly.
sim_sum_1 <- matrix(data = NA, nrow = TMAX, ncol = 2);
colnames(sim_sum_1) <- c("Time", "Pop_size");
moving_month <- 0;

for(time_step in 1:TMAX){
  sim_new                 <- gmse_apply(old_list = sim_old, get_res = 'Full',
                                        res_move_type = moving_month);
  sim_sum_1[time_step, 1] <- time_step;
  sim_sum_1[time_step, 2] <- sim_new$basic_output$resource_results[1];
  
  next_time               <- time_step + 1; # What happens next time step?
  if(next_time %in% JAN){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.976083968; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  if(next_time %in% FEB){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.94824969; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  if(next_time %in% MAR){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.9023544; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  if(next_time %in% APR){
    moving_month                      <- 1; # A moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.811569975; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  if(next_time %in% MAY){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.805694059; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
  }
  if(next_time %in% JUN){
    moving_month                      <- 1; # A moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.794513672; # Death probability
    sim_new[["resource_array"]][, 10] <- 3;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  if(next_time %in% JUL){
    moving_month                      <- 1; # A moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.767700722; # Death probability
    sim_new[["resource_array"]][, 10] <- 3;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
  }
  if(next_time %in% AUG){
    moving_month                      <- 1; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.842305069; # Death probability
    sim_new[["resource_array"]][, 10] <- 3;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
  }
  if(next_time %in% SEP){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.877005333; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
  }
  
  if(next_time %in% OCT){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.891729601; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  if(next_time %in% NOV){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.939895976; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  if(next_time %in% DEC){
    moving_month                      <- 0; # Not a moving month
    sim_new[["resource_array"]][, 9]  <- 1 - 0.956833861; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
  }
  sim_old <- sim_new; # BD: This should always go at the end
  
  #print(sim_sum_1[time_step,]); #Activate to see the simulation progress
}
  
}, error = function(e) { #tryCatch: if error occurs because N = 0
    cat("Error occurred in iteration", rep, ":", conditionMessage(e), "\n")
}) #end of tryCatch


sim_sum_1 <- cbind(rep, sim_sum_1); # Add iteration number as a column
print(sim_sum_1); 
results[[rep]] <- sim_sum_1

}

combined_results <- do.call(rbind, results)
combined_results[is.na(combined_results)] <- 0

#now timestep is also recorded as a 0 from the moment extinction occurred, we want time step to be recorded correctly

combined_results <- as.data.frame(combined_results)

# Loop through each repetition
for (rep_id in unique(combined_results$rep)) {
  # Subset data for the current repetition
  rep_data <- combined_results[combined_results$rep == rep_id, ]
  
  # Loop through each row in the current repetition's data
  for (i in 2:nrow(rep_data)) {
    # If the current time step is 0, replace it with the previous time step + 1
    if (rep_data$Time[i] == 0) {
      rep_data$Time[i] <- rep_data$Time[i - 1] + 1
    }
  }
  
  # Replace the data for the current repetition in the original data frame
  combined_results[combined_results$rep == rep_id, ] <- rep_data
}

file_path <- "C:/Users/imket/Documents/GitHub/European_hamster_GMSE/Output/Simulation8_100it_250startN.csv" #Give the simulation an appropriate name
write.csv(combined_results, file = file_path, row.names = FALSE)

```


## 2: Simulations modelling multiple different crop types on the landscape. Crop type influences the population parameters of the hamsters ocurring in it.

```{r, , echo=TRUE, eval=FALSE}
library(GMSE)

#parameters specified below are for hamsters in crop type 2. For type 3 an extra litter is modelled, for type 1 a different mortality is modelled (no survival) 

DIM_1 <- 447; # Land dimension 1
DIM_2 <- 447; # Land dimension 2

iterations <- 100 #number of iterations
results <- list() #list to store results from each repetition

for (rep in 1:iterations) {
  tryCatch({
    
################################### Adding the landscape function here = different landscape for each iteration   
crop_type2 <- function(land, agents = 11) { # agents = 1 (manager) + number of stakeholders
  xdim <- dim(land)[1]
  ydim <- dim(land)[2]
  
  # Randomly assign stakeholders to crop types
  stakeholders_vector <- sample(2:agents) #randomly shuffle the stakeholders (agents), but leave out the first (= always the manager)
  type3_stakeholders <- stakeholders_vector[1:2] #2 get type 3
  type2_stakeholders <- stakeholders_vector[3:6] #4 get type 2
  type1_stakeholders <- stakeholders_vector[7:10] #4 get type 1
  
  # Create a vector to store the crop type for each stakeholder
  stakeholder_types <- rep(NA, agents)
  
  for (stakeholder in 2:agents) { 
    if (stakeholder %in% type3_stakeholders) {
      stakeholder_types[stakeholder] <- 3
    } else if (stakeholder %in% type2_stakeholders) {
      stakeholder_types[stakeholder] <- 2
    } else if (stakeholder %in% type1_stakeholders) {
      stakeholder_types[stakeholder] <- 1
    }
  }

  for (i in 1:xdim) {   
    for (j in 1:ydim) {
      if (land[i, j, 3] %in% 2:agents) {
        land[i, j, 1] <- stakeholder_types[land[i, j, 3]]
      }
    } 
  }
  return(land)
}
#####################################################

# Initialise the first output
sim_old   <- gmse_apply(stakeholders    = 10, # should be the same as in function crop_type!
                        res_movement    = 100, 
                        res_move_type   = 0,
                        remove_pr       = 1 - 0.976083968, 
                        lambda          = 0, 
                        res_death_type  = 1,
                        observe_type    = 2, # but only once a year
                        res_move_obs    = FALSE, 
                        max_ages        = 24, 
                        RESOURCE_ini    = 250, # start N 
                        culling         = FALSE,
                        land_ownership  = TRUE,
                        age_repr        = 4,
                        land_dim_1      = DIM_1,
                        land_dim_2      = DIM_2,
                        manage_target   = 2500,
                        get_res         = 'Full');
old_obs  <- sim_old[["observation_array"]];

# Add crop types on layer 1 of the landscape with the crop_type(2) function
sim_old[["LAND"]] <- crop_type2(land = sim_old[["LAND"]]);

# Initialise the results matrix for the current iteration; Currently, 3 variables are recorded as output
# Double-brackets: technically a bit more secure. It works if you edit columns of the resource_array directly.
sim_sum_1 <- matrix(data = NA, nrow = TMAX, ncol = 3);
colnames(sim_sum_1) <- c("Time", "Pop_size", "Pop_est")
moving_month <- 0;

for(time_step in 1:TMAX){
  sim_new                 <- gmse_apply(old_list = sim_old, get_res = 'Full',
                                        res_move_type = moving_month);
  sim_sum_1[time_step, 1] <- time_step;
  sim_sum_1[time_step, 2] <- sim_new$basic_output$resource_results[1];
  sim_sum_1[time_step, 3] <- sim_new$basic_output$observation_results[1];
  
  next_time               <- time_step + 1; # What happens next time step?
  
  if(next_time %in% JAN){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.976083968; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% FEB){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.94824969; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% MAR){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.9023544; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% APR){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.811569975; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% MAY){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.805694059; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% JUN){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.794513672; # Death probability
    sim_new[["resource_array"]][, 10] <- 1.19;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% JUL){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.767700722; # Death probability
    sim_new[["resource_array"]][, 10] <- 1.19;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                      res  = temp_res);
  }
  if(next_time %in% AUG){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.842305069; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land,  
                                                        res  = temp_res)
  }
  if(next_time %in% SEP){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.877005333; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);

    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  
  if(next_time %in% OCT){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.891729601; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
    
    
  }
  if(next_time %in% NOV){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.939895976; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% DEC){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.956833861; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  sim_old <- sim_new; # This should always go at the end
  
  #print(sim_sum_1[time_step,]); #Activate to see the simulation progress
}

}, error = function(e) { #tryCatch: if error occurs because N = 0
    cat("Error occurred in iteration", rep, ":", conditionMessage(e), "\n")
}) #end of tryCatch

sim_sum_1 <- cbind(rep, sim_sum_1); # Add iteration number as a column
print(sim_sum_1); 
results[[rep]] <- sim_sum_1

###################################### Extra function to save landscape layout for each simulation
# Get the current timestamp
timestamp <- format(Sys.time(), "%Y%m%d-%H%M%S")

# Crop type plot
# Adjust plot margins to make space for the legend
output_folder<- "C:/Users/imket/Documents/UAntwerpen/Master thesis/Figures/Landscape simulations/Simulation 8"
filename_crop <- file.path(output_folder, paste0(timestamp, ".png"))
png(filename_crop, width = 1000, height = 800)
par(mar = c(5, 4, 4, 8) + 0.1)
image(z = sim_old$LAND[,,1], col = c("#FF6961", "#EBD657", "#92C37F"), xaxt = "n", yaxt = "n", 
      main = "Crop Type Distribution", bty = "n", cex.main = 1.5, cex.lab = 1.2)
legend("right", inset = c(-0.13, 0),  legend = c("Crop Type 1", "Crop Type 2", "Crop Type 3"), 
       fill = c("#FF6961", "#EBD657", "#92C37F"), xpd = TRUE, cex = 1) # to show crop type distribution in the landscape
dev.off()
#####################################

}

combined_results <- do.call(rbind, results)
combined_results[is.na(combined_results)] <- 0

#now timestep is also recorded as a 0 from the moment extinction occurred, we want time step to be recorded correctly

combined_results <- as.data.frame(combined_results)

# Loop through each repetition
for (rep_id in unique(combined_results$rep)) {
  # Subset data for the current repetition
  rep_data <- combined_results[combined_results$rep == rep_id, ]
  
  # Loop through each row in the current repetition's data
  for (i in 2:nrow(rep_data)) {
    # If the current time step is 0, replace it with the previous time step + 1
    if (rep_data$Time[i] == 0) {
      rep_data$Time[i] <- rep_data$Time[i - 1] + 1
    }
  }
  
  # Replace the data for the current repetition in the original data frame
  combined_results[combined_results$rep == rep_id, ] <- rep_data
}

file_path <- "C:/Users/imket/Documents/GitHub/European_hamster_GMSE/Output/Simulation8_100it_250startN.csv" #Give the simulation an appropriate name
write.csv(combined_results, file = file_path, row.names = FALSE)

```

## 3 - EXTRA NOV2024: Simulations modelling multiple different crop types on the landscape. Crop type influences the population parameters of the hamsters ocurring in it.

```{r, , echo=TRUE, eval=FALSE}
library(GMSE)

#parameters specified below are for hamsters in crop type 2. For type 3 an extra litter is modelled, for type 1 a different mortality is modelled (no survival) 

DIM_1 <- 447; # Land dimension 1
DIM_2 <- 447; # Land dimension 2

iterations <- 100 #number of iterations
results <- list() #list to store results from each repetition

for (rep in 1:iterations) {
  tryCatch({
    
################################### Adding the landscape function here = different landscape for each iteration   
crop_type2 <- function(land, agents = 101) { # agents = 1 (manager) + number of stakeholders
  xdim <- dim(land)[1]
  ydim <- dim(land)[2]
  
  # Randomly assign stakeholders to crop types
  stakeholders_vector <- sample(2:agents) #randomly shuffle the stakeholders (agents), but leave out the first (= always the manager)
  type3_stakeholders <- stakeholders_vector[1:20] #2 get type 3
  type2_stakeholders <- stakeholders_vector[21:60] #4 get type 2
  type1_stakeholders <- stakeholders_vector[61:100] #4 get type 1
  
  # Create a vector to store the crop type for each stakeholder
  stakeholder_types <- rep(NA, agents)
  
  for (stakeholder in 2:agents) { 
    if (stakeholder %in% type3_stakeholders) {
      stakeholder_types[stakeholder] <- 3
    } else if (stakeholder %in% type2_stakeholders) {
      stakeholder_types[stakeholder] <- 2
    } else if (stakeholder %in% type1_stakeholders) {
      stakeholder_types[stakeholder] <- 1
    }
  }

  for (i in 1:xdim) {   
    for (j in 1:ydim) {
      if (land[i, j, 3] %in% 2:agents) {
        land[i, j, 1] <- stakeholder_types[land[i, j, 3]]
      }
    } 
  }
  return(land)
}
#####################################################

# Initialise the first output
sim_old   <- gmse_apply(stakeholders    = 100, 
                        res_movement    = 100, 
                        res_move_type   = 0,
                        remove_pr       = 1 - 0.976083968, 
                        lambda          = 0, 
                        res_death_type  = 1,
                        observe_type    = 2, 
                        res_move_obs    = FALSE, 
                        max_ages        = 24, 
                        RESOURCE_ini    = 250, # start N 
                        culling         = FALSE,
                        land_ownership  = TRUE,
                        age_repr        = 4,
                        land_dim_1      = DIM_1,
                        land_dim_2      = DIM_2,
                        manage_target   = 2500,
                        get_res         = 'Full');
old_obs  <- sim_old[["observation_array"]];

# Add crop types on layer 1 of the landscape with the crop_type(2) function
sim_old[["LAND"]] <- crop_type2(land = sim_old[["LAND"]]);

# Initialise the results matrix for the current iteration; Currently, 3 variables are recorded as output
# Double-brackets: technically a bit more secure. It works if you edit columns of the resource_array directly.
sim_sum_1 <- matrix(data = NA, nrow = TMAX, ncol = 2);
colnames(sim_sum_1) <- c("Time", "Pop_size")
moving_month <- 0;

for(time_step in 1:TMAX){
  sim_new                 <- gmse_apply(old_list = sim_old, get_res = 'Full',
                                        res_move_type = moving_month);
  sim_sum_1[time_step, 1] <- time_step;
  sim_sum_1[time_step, 2] <- sim_new$basic_output$resource_results[1];
  
  next_time               <- time_step + 1; # What happens next time step?
  
  if(next_time %in% JAN){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.976083968; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% FEB){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.94824969; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% MAR){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.9023544; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% APR){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.811569975; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% MAY){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.805694059; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% JUN){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.794513672; # Death probability
    sim_new[["resource_array"]][, 10] <- 1.19;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% JUL){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.767700722; # Death probability
    sim_new[["resource_array"]][, 10] <- 1.19;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                      res  = temp_res);
  }
  if(next_time %in% AUG){
    moving_month                      <- 1;         # Movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.842305069; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land,  
                                                        res  = temp_res)
  }
  if(next_time %in% SEP){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.877005333; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    temp_res                          <- sim_new[["resource_array"]];
    sim_new[["resource_array"]]       <- cell_K(res   = temp_res, DIM_1 = DIM_1, 
                                                DIM_2 = DIM_2);

    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  
  if(next_time %in% OCT){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.891729601; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
    
    
  }
  if(next_time %in% NOV){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.939895976; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  if(next_time %in% DEC){
    moving_month                      <- 0;         # No movement 
    sim_new[["resource_array"]][, 9]  <- 1 - 0.956833861; # Death probability
    sim_new[["resource_array"]][, 10] <- 0;         # Birth probability
    sim_new[["observation_array"]]    <- old_obs;   # Use old observations
    
    temp_res                          <- sim_new[["resource_array"]];
    temp_land                         <- sim_new[["LAND"]];
    sim_new[["resource_array"]]       <- crop_hamster(land = temp_land, 
                                                        res  = temp_res);
  }
  sim_old <- sim_new; # This should always go at the end
  
  #print(sim_sum_1[time_step,]); #Activate to see the simulation progress
}

}, error = function(e) { #tryCatch: if error occurs because N = 0
    cat("Error occurred in iteration", rep, ":", conditionMessage(e), "\n")
}) #end of tryCatch

sim_sum_1 <- cbind(rep, sim_sum_1); # Add iteration number as a column
print(sim_sum_1); 
results[[rep]] <- sim_sum_1

###################################### Extra function to save landscape layout for each simulation
# Get the current timestamp
timestamp <- format(Sys.time(), "%Y%m%d-%H%M%S")

# Crop type plot
# Adjust plot margins to make space for the legend
output_folder<- "C:/Users/lucp13135/Documents/Simulations/Simulation 11b"
filename_crop <- file.path(output_folder, paste0(timestamp, ".png"))
png(filename_crop, width = 1000, height = 800)
par(mar = c(5, 4, 4, 8) + 0.1)
image(z = sim_old$LAND[,,1], col = c("#FF6961", "#EBD657", "#92C37F"), xaxt = "n", yaxt = "n", 
      main = "Crop Type Distribution", bty = "n", cex.main = 1.5, cex.lab = 1.2)
legend("right", inset = c(-0.13, 0),  legend = c("Crop Type 1", "Crop Type 2", "Crop Type 3"), 
       fill = c("#FF6961", "#EBD657", "#92C37F"), xpd = TRUE, cex = 1) # to show crop type distribution in the landscape
dev.off()
#####################################
}

combined_results <- do.call(rbind, results)
combined_results[is.na(combined_results)] <- 0

#now timestep is also recorded as a 0 from the moment extinction occurred, we want time step to be recorded correctly

combined_results <- as.data.frame(combined_results)

# Loop through each repetition
for (rep_id in unique(combined_results$rep)) {
  # Subset data for the current repetition
  rep_data <- combined_results[combined_results$rep == rep_id, ]
  
  # Loop through each row in the current repetition's data
  for (i in 2:nrow(rep_data)) {
    # If the current time step is 0, replace it with the previous time step + 1
    if (rep_data$Time[i] == 0) {
      rep_data$Time[i] <- rep_data$Time[i - 1] + 1
    }
  }
  
  # Replace the data for the current repetition in the original data frame
  combined_results[combined_results$rep == rep_id, ] <- rep_data
}

file_path <- "C:/Users/lucp13135/Documents/Simulations/Simulation 11b/Simulation11b_100it_250startN.csv" #Give the simulation an appropriate name
write.csv(combined_results, file = file_path, row.names = FALSE)

```



```{r, , echo=TRUE, eval=FALSE}
# Crop type plot
# Adjust plot margins to make space for the legend
par(mar = c(2, 3, 2, 8) + 0.2)
image(z = sim_old$LAND[,,1], col = c("#FF6961", "#F8D66D", "#8CD47E"), xaxt = "n", yaxt = "n", main = "Crop type distribution", bty = "n")
legend("right", inset = c(-0.30, 0),  legend = c("Crop Type 1", "Crop Type 2", "Crop Type 3"), 
fill = c("#FF6961", "#F8D66D", "#8CD47E"), xpd = TRUE, cex = 1) #to show crop type distribution in the landscape

# Stakeholders plot 
# Define the range of stakeholder numbers
stholders <- 2:11 #adjust based on number of stakeholders 

# Generate a color palette with enough distinct colors for each stakeholder
colours <- rainbow(length(stholders), s = 0.5)
image(z = sim_old$LAND[,,3], col = colours, xaxt = "n", yaxt = "n", main = "Stakeholder land distribution", bty = "n")

```

